---
title: "course_analysis"
author: "Rica Rebusit"
date: "4/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(Hmisc) # Enables %nin% notation for "not in"
library(readxl) # read Excel file
library(magrittr) #Allows %<>% notation to update lhs object with resulting value
library(naniar)
library(tinytex)
library(ggplot2)
library(kableExtra)
library(broom)
library(dotwhisker)
library(dplyr)
library(ggthemes)
library(stargazer)

# Run everything from the data_management.R file
course.level <- read_csv("data/course_analysis_dataset.csv")
# Convert to factors
course.level$Term.Year<-factor(course.level$Term.Year)
course.level$SI.Component.Flag<-factor(course.level$SI.Component.Flag)
No.SI <- filter(course.level, SI.Component.Flag == "0") # Looking at the courses less than 10 SI visits
```

# Initial Graphs

```{r}
ggplot(course.level, aes(URM, dwf.rate, color = URM)) + geom_point() + facet_wrap(. ~Term.Year) + xlab("Percent URM") + ylab("DWF Rate") + geom_point(data=No.SI, aes(x=URM, y=dwf.rate), color = 'red') + theme(legend.position="none") + geom_smooth(se=F) 
```
It seems like over the years, if the percent of URM increases in a course, the dwf.rate decreases. Possibly be because URM "works" harder in a course that offers SI?

```{r}
ggplot(course.level, aes(First.Gen.Perc, dwf.rate, color = First.Gen.Perc)) + geom_point() + facet_wrap(. ~Term.Year) + xlab("Percent First Gen") + ylab("DWF Rate")+ geom_point(data=No.SI, aes(x=URM, y=dwf.rate), color = 'red') + theme(legend.position="none") + geom_smooth(se=F)
```
Same thing with the percentage of first generation students in a course. As the percentage increases the dwf rates decreases. So a course that offers SI helps courses with a high percentage of equity gaps decrease dwf rates


```{r}
ggplot(course.level, aes(Term.Year, dwf.rate, color = URM)) + 
  geom_line() +
  scale_y_continuous(limits = c(0, max(course.level$dwf.rate)))
```

```{r}
ggplot(course.level, aes(Term.Year, dwf.rate, color = First.Gen.Perc)) + 
  geom_line() +
  scale_y_continuous(limits = c(0, max(course.level$dwf.rate)))
```

# Modelling
```{r}
lm6 <- lm(dwf.rate ~ SI.Component.Flag + URM + First.Gen.Perc + Term.Year + HS.GPA.Ave + class.average, data = course.level)
summary(lm6)
```

Term Year 2018 and 2019 are statistically significant as well as class average

```{r}
lm3 <- lm(dwf.rate ~ SI.Component.Flag + URM + First.Gen.Perc + Term.Year + HS.GPA.Ave + URM*SI.Component.Flag + First.Gen.Perc*SI.Component.Flag, data=course.level, na.action = na.omit)
summary(lm3)
```

First Generation is statistically significant

```{r}
lm2 <- lm(dwf.rate ~ SI.Component.Flag + URM + First.Gen.Perc + Term.Year, data = course.level)
summary(lm2)
model2 <- tidy(lm2)
dwplot(model2) + geom_vline(xintercept = 0, colour = "black",linetype = 2) + ggtitle("Model Results DWF Rate and Equity Gaps") + xlab("Confidence Interval") + theme_few() +
  theme(legend.position = "none", axis.title.x = element_text(size = 20), title = element_text(size = 32))
d2 <- data.frame(model2)
kable(d2)
```

# Stargazer

```{r}
m <- lm(dwf.rate ~ SI.Component.Flag + URM + First.Gen.Perc + class.size + class.average, data = course.level)
summary(m)
mod <- tidy(m)
df <- data.frame(mod)
kable(mod)
```

```{r}
m2 <- lm(dwf.rate ~ SI.Component.Flag + URM + First.Gen.Perc + SI.Component.Flag*URM + SI.Component.Flag*First.Gen.Perc + class.size + class.average, data = course.level)
summary(m2)
mod2 <- tidy(m2)
df2 <- data.frame(mod2)
kable(mod)
```

```{r, results = 'asis', echo = FALSE}
stargazer(m, m2, off.ratio=T, single.row = TRUE, type = "html", title = "Results Multilinear Regression")
```

