---
title: Exploring the Effect of Supplimental Instruction on Equity Gaps 
author:
  - name: Faith Fatchen
    affil: 1
  - name: Skip Moses
    affil: 1
  - name: Rica Rebusit
    affil: 1
  - name: Joseph Shifman
    affil: 1
  - name: Robin Donatello
    affil: 1
  - name: Eric Wasigner
    affil: 2
affiliation:
  - num: 1
    address: Department of Mathematics and Statistics, California State University, Chico
  - num: 2
    address: Department of Chemistry and Biochemistry, California State University, Chico
column_numbers: 3
primary_colour: "#9D2235"
secondary_colour: "#FFFFFF"
poster_height: "36in"
poster_width: "60in"
output: 
  posterdown::posterdown_html:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(tidyverse)
library(readxl)
library(magrittr) #Allows %<>% notation to update lhs object with resulting value
library(naniar)
library(Hmisc) # %nin%
library(stargazer)
library(tinytex)
library(ggplot2)
library(dotwhisker)
library(broom)
library(multcomp)
library(ggthemes)

d<-read_csv("~/Documents/consulting-project-SI/data/student_analysis_dataset.csv")
d[sapply(d, is.character)] <- lapply(d[sapply(d, is.character)], as.factor)
d$Term.Year<-factor(d$Term.Year)
d$Cohort.Term.Year<-factor(d$Cohort.Term.Year)
d<-rename(d, URM = IPEDS.Ethnicity.URM.Non.URM)
d$URM.First.Gen<-factor(ifelse(d$URM=="URM" & d$First.Generation.Flag == "Y", 1, 0 )) #new var, student who is urm and first gen
d$One.Year.Retention<-factor(d$One.Year.Retention)
fit1<-glm(One.Year.Retention ~ Cohort.Term.Year + SI.Attended + Student.Class.Unit.Attempted + Student.Class.Unit.Passed +  URM + Gender.Code + First.Generation.Flag + Academic.Level + Major.1.STEM.Flag + Academic.Standing.Status + HS.GPA + Inst.MD.Persn.Onl.Othr + Course.Fee.Exist.Flag + GE.Class.Flag, family = binomial(link = "logit"), data = d)

fit3<-glm(One.Year.Retention~ Cohort.Term.Year + SI.Attended + Student.Class.Unit.Attempted + Student.Class.Unit.Passed +  URM + Gender.Code + First.Generation.Flag + Academic.Level + Major.1.STEM.Flag + Academic.Standing.Status + HS.GPA + Inst.MD.Persn.Onl.Othr + Course.Fee.Exist.Flag + GE.Class.Flag + SI.Attended:URM +SI.Attended:First.Generation.Flag, family = binomial(link = "logit"), data = d)

dod_data <- d %>% group_by(URM, SI.Attended)
dod_data$One.Year.Retention <- factor(dod_data$One.Year.Retention, labels = c("No", "Yes"))
dod_data2<-d %>% group_by(First.Generation.Flag, SI.Attended)
dod_data2$One.Year.Retention <- factor(dod_data2$One.Year.Retention, labels = c("No", "Yes"))

course.level <- read_csv("data/course_analysis_dataset.csv")
dod <- course.level %>% group_by(URM, SI.Component.Flag)
```
# Introduction

<<<<<<< HEAD
#### What is Supplemental Instruction? 
=======
### What is Supplemental Instruction? 
>>>>>>> bac85ce1b463b02bf6dfafe8a4024a9697728440

* Supplemental Instruction (SI) is designed to increase student success in historically difficult courses by utilizing peer lead study session. The SI leader has performed well in the course. Using this experience, the SI leader encourages and assists students own success. Students success is a top mission throughtout universities. Hence, it is of the utmost important to find and/or develop strategies to promote student success in the classroom. Students come from diverse backgrounds, and may need additionaly support to flourish at the university level. SI aims bridge this gap, so every student has an equal opportunity at success. 

#### Related Work and Background (Rica could put her summaries here)

* The impact of SI can be addressed at the class level or the individual student level. 
Here, we investigate both. We obtained data from CSU Chico Institutional Research. 
The data included course information, student demographics, and student course grades. We considered measures of success at both the student level and course. The methods for these different explorations are detailed below.

#### What is an Equity Gap?
* By measuring equity gaps, we seek to address disparities faced by historically underserved groups. When addressing inequities, CSU's consider three student groups underrepresented minorities (URM), first generation, and low income. Due to data limitations, here we consider the former two groups. 
  
# Methods

  <table>
<caption>Table 1: Population and Sample Comparison</caption>
 <thead>
  <tr>
   <th style="text-align:left;"> Variable </th>
   <th style="text-align:right;"> Chancelor.Data </th>
   <th style="text-align:right;"> Sample.Data </th>
  </tr>
 </thead>
<tbody>
  <tr>
   <td style="text-align:left;"> Mean Number of Students </td>
   <td style="text-align:right;"> 16984 </td>
   <td style="text-align:right;"> 16075 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Mean Percentage of Females </td>
   <td style="text-align:right;"> 54.27 </td>
   <td style="text-align:right;"> 53.34 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Mean Percentage of Males </td>
   <td style="text-align:right;"> 45.72 </td>
   <td style="text-align:right;"> 46.61 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Mean Percentage of First Gen. </td>
   <td style="text-align:right;"> 50.94 </td>
   <td style="text-align:right;"> 38.13 </td>
  </tr>
  <tr>
   <td style="text-align:left;"> Mean Percentage of URM </td>
   <td style="text-align:right;"> 36.66 </td>
   <td style="text-align:right;"> 37.55 </td>
  </tr>
</tbody>
</table>


### Student Level

*  The student level explores the success of an individual student. We define student success as one year retention. This means the student was still enrolled one year from the course. At the student level, the questions we are addressing is "Does SI have an impact on the one retention". Then, to consider the the measures of equity previously defined, "Does the impact of SI differ between the equity groups".
We limited our sample to first time freshman. Hence, they should still be enrolled one year later.
Additionally, is primarily found in GE classes, so transfer students are likely not enrolling in 
these course at CSU Chico. Figure 1 presents one year retention for those who did and did not attend SI by one of the measures of equity, URM. 

<<<<<<< HEAD
```{r echo=FALSE}
=======
```{r echo=FALSE, fig.dim = c(17,7)}
>>>>>>> bac85ce1b463b02bf6dfafe8a4024a9697728440
facet_names<-c(`0` = "Did Not Attend SI", `1` = "Attended SI")
ggplot(dod_data) +
  geom_bar(aes(x=One.Year.Retention, fill=URM),
           position = "dodge") +
  facet_wrap(~SI.Attended, labeller = as_labeller(facet_names)) + ggtitle("Figure 1") +xlab("One Year Retention")+ scale_fill_discrete(name="Underrepresented\nMinority(URM)") + ylab("") + theme_few() +
  theme(text = element_text(size = 32))

ggplot(dod_data2) +
  geom_bar(aes(x=One.Year.Retention, fill=First.Generation.Flag),
           position = "dodge") +
  facet_wrap(~SI.Attended, labeller = as_labeller(facet_names)) + ggtitle("Figure 2") +xlab("One Year Retention")+ scale_fill_discrete(name="First\nGeneration") + ylab("") + theme_few() +
  theme(text = element_text(size = 32))
```


#### Course Level

<<<<<<< HEAD
```{r}
ggplot(dod, aes(dwf.rate, color=URM)) + geom_density(alpha = 0.2) + theme_bw(base_size = 10) + xlab("DWF Rate")
```

```{r}
=======
```{r eval=FALSE, include=FALSE}
#Does not work
ggplot(dod, aes(dwf.rate, color=URM)) + geom_density(alpha = 0.2) + theme_bw(base_size = 10) + xlab("DWF Rate")
```

```{r eval=FALSE, include=FALSE}
#does not work
>>>>>>> bac85ce1b463b02bf6dfafe8a4024a9697728440
ggplot(dod, aes(SI.Component.Flag, dwf.rate, color=URM)) + geom_boxplot() + xlab("Attended SI") + ylab("DWF Rate")
```


# Student Level Models
  
```{r echo=FALSE, fig.width=17}
m1<-tidy(fit1) %>%
  filter(term == "SI.Attended" | term == "URMURM" | term == "First.Generation.FlagY")
dwplot(m1) +
    geom_vline(xintercept = 0, colour = "black",linetype = 2) + ggtitle("Figure 3: Model Results SI and Equity Measures") + xlab("Confidence Interval") + theme_few() +
  theme(legend.position = "none", axis.title.x = element_text(size = 20), title = element_text(size = 32))
m4<-tidy(fit3)%>%
  filter(term == "SI.Attended" | term == "URMURM" | term == "First.Generation.FlagY" | term == "SI.Attended:URMURM" | term == "SI.Attended:First.Generation.FlagY")
dwplot(m4) +
    geom_vline(xintercept = 0, colour = "grey60", linetype = 2) + ggtitle("Figure 4: Model Results Interactions SI and Equity Measures")+ xlab("Confidence Interval") + theme_few() +
  theme(legend.position = "none", axis.title.x = element_text(size = 20), title = element_text(size = 32))
```

* Figure 3, presents the results from a model considering the first student level question, "Does SI have an impact on the one retention". The results from this model imply, that SI does not a have a significant impact on student one year retention. If the student is first generation is also insignificant. Additionally, this model also implies that a student that is classified as URM is positively related to one year retention. Figure 4, presents the results from a model considering the second student level question "Does the impact of SI differ between the equity groups". Again, SI attendance was found to be insignificant and URM is positively related. Additionally, if the student is an URM and attended SI, was also found to be insignificant. If the student is first generation is also found to be insignificant. However, if the student is first generation and attends SI the probability of year retention increases, and this relationship is significant. For both models, other covariates that may capture variation in one year retention were included in the model that are not presented here. 
* 
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *

# Course Level Models 
  
```{r}
ggplot(course.level, aes(dwf.rate, class.average, color=URM)) + geom_point(alpha=0.1) + geom_smooth(method=lm)
```
  
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
<<<<<<< HEAD
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
  *
=======
# Conclusions
* The results from these initial models did not find significant results for our variables of interest. However, this could be due to underlying causes such as omitted variable bias. The following section describes additional work that may help us better consider our questions. 
>>>>>>> bac85ce1b463b02bf6dfafe8a4024a9697728440

# Future Work

  * Ideally, we want to identify attending SI as a treatment, and explore it's causal effect on equity gaps.
  * In order to accomplish this, we will employ Coarsened Exact Matching to control for counfounding variables. 
  * With this matching we can estimate the effect of SI on equity gaps with SI as a treatment as in a clinical trial. 
  
# Conclusions
