---
title: "Untitled"
author: "Skip Moses"
date: "4/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(MatchIt)
library(lmtest)
library(sandwich)
library(ggpubr)

library(multcomp) # difference of differnces
library(scales) # plotting
```


```{r}
# 2016-2019 Student grades/SI/demographic data
data_raw <- read.csv("data/CEM_full_dataset.csv")
#data_raw <- read.csv("C:/Box/Projects_Current/SI Equity analysis/data/CEM_full_dataset.csv")

data <- data_raw %>% filter(Term.Year < 2020, Entry.Enrollment.Type == "First-Time Freshman")  %>%
  # Remove Withdraws from class 
  filter(Student.Class.Official.Grade != "W") %>%
  filter(Student.Class.Official.Grade != "WU") %>%
  filter(Student.Class.Official.Grade != "IC") %>%
  filter(Student.Class.Official.Grade != "AU")

# COnvert to factors
data$SI.Attended <- as.factor(data$SI.Attended)
data$Random.Student.ID <- as.factor(data$Random.Student.ID)
data$Random.Course.ID <- as.factor(data$Random.Course.ID)

# Center the grade per unit (this is actually normalizing - RAD)
# And it's normalizing against the mean/sd for the entire data set. 
# normalizing within class would be more appropriate. 
# (IF we really need to go there. let's keep the normal grade scale for now.)
#data$Grade.Point.per.Unit.center <- (data$Student.Class.Grade.Point.per.Unit - mean(data$Student.Class.Grade.Point.per.Unit))/sd(data$Student.Class.Grade.Point.per.Unit)
```


```{r}
# Turning this off to to calculate glm derived propensity
# Estimate Propensity Scores
prop.score <- glm(SI.Attended ~ HS.GPA + 
                    Student.Orientation.Flag + 
                    Major.1.STEM.Flag +
                    Random.Course.ID, family = binomial, data = data)

summary(prop.score)

data$Prop.Score <- predict(prop.score, type = "response")

labs <- paste("Actually Attended SI:", c("Did not Attend", "Attended"))

# Histogram of Propensity Scores (I think the labels are reversed)
#data %>% mutate(SI.Attended = ifelse(SI.Attended == 1, labs[1], labs[2])) %>% 
#  ggplot(aes(x = Prop.Score)) +
#    geom_histogram(binwidth = .1) + facet_wrap(~SI.Attended) + xlab("Probability of goint to SI")

data %>% mutate(SI.Attended = ifelse(SI.Attended == 1, labs[2], labs[1])) %>% 
  ggplot(aes(x = Prop.Score, y=First.Generation.Flag, fill=SI.Attended)) +
    geom_violin() + 
  facet_wrap(~IPEDS.Ethnicity.URM.Non.URM) + coord_flip()+
  xlab("Probability of goint to SI")
```


```{r}
# Here we make our matching. The k2K = TRUE will coarsen the covariates first
# and then do an exact matching on the coarsend data. This allows us to compute 
# the difference in grade between our matching.
matching <- matchit(SI.Attended ~ 
                      HS.GPA + 
                      Student.Orientation.Flag + 
                      Major.1.STEM.Flag +
                      Random.Course.ID ,
                    data = data, 
                    method = 'cem', estimand = 'ATE',
                    k2k = TRUE, # After Pair Matching without Replacement
                    k2k.method = "euclidean")
matching
#summary(matching)
```


```{r}
# Covariates should be those that cause variation in the outcome and selection into treatment group; these are known as confounding variables. (Greifer)
matching.strict <- matchit(SI.Attended ~ 
                      HS.GPA + 
                      Student.Orientation.Flag + 
                      Major.1.STEM.Flag +
                      Random.Course.ID + 
                      Academic.Level + 
                      First.Generation.Flag + 
                      IPEDS.Ethnicity.URM.Non.URM +
                      Gender.Code, 
                    data = data, 
                    method = 'cem', estimand = 'ATT',
                    k2k = TRUE, # FALSE - After Stratum Matching
                    k2k.method = "euclidean")
matching.strict

# Extract Matched pairs
matched_df1 <- match.data(matching) %>% 
  relocate(subclass, SI.Attended) %>% 
  arrange(subclass, SI.Attended)  %>%
  rename(grade = Student.Class.Grade.Point.per.Unit)
```


## Exploratory plots with matched sample - trying to create this gap
```{r, fig.width=10}
# Identify courses with 20 or more students (10 matched pairs per class)
matched_courses <- matched_df1 %>% group_by(Random.Course.ID) %>%
  tally() %>% filter(n>=20) %>% droplevels()

matched_df2 <- matched_df1 %>% 
  filter(Random.Course.ID %in% matched_courses$Random.Course.ID) %>%
  group_by(Random.Course.ID, IPEDS.Ethnicity.URM.Non.URM, SI.Attended,
           Term.Year, Term.Type, Academic.Subject, Class.Subject.Number.Section) %>%
  summarise(avg.grade = mean(grade), 
            n=n()) 

matched_df2_wide <- matched_df2 %>% 
  pivot_wider(id_cols = c(Random.Course.ID, SI.Attended, Term.Year, Term.Type, Academic.Subject, Class.Subject.Number.Section), 
              names_from = IPEDS.Ethnicity.URM.Non.URM, 
              values_from = c(avg.grade, n)) %>%
  mutate(urm_gap = `avg.grade_Non-URM` - avg.grade_URM, 
         n = pmin(`n_Non-URM`, n_URM)) 

ggplot(matched_df2_wide, 
       aes(y=urm_gap, x=SI.Attended, color = Random.Course.ID, 
           group = Random.Course.ID)) + 
  geom_point(aes(size=n)) + geom_line()  + geom_hline(yintercept=0) + 
  geom_text(aes(x=.75, y=.25, label = "Non-URM > URM"), col="red") +
  geom_text(aes(x=.75, y=-.25, label = "Non-URM < URM"), col="green3") +
  ylab("Non-FG - FG") + 
  ggtitle("URM gap among matched students who went to SI \n vs those who did not go to SI")

matched_df2_wide %>% filter(Term.Year > 2016)  %>%
ggplot(aes(y=urm_gap, x=SI.Attended, color = Academic.Subject, 
           group = Random.Course.ID)) + 
  geom_point(aes(size=n)) + geom_line()  + geom_hline(yintercept=0) +
  ylab("Non-URM - URM") + xlab("Attended SI")  + guides(size = "none") + labs(color = "Subject") +
  ggtitle("URM gap among matched students who went to SI\n vs those who did not go to SI") + 
  facet_wrap(~Term.Year + Term.Type) 
ggsave("poster/pictures/urm_gap_by_year.png")
```


```{r, fig.width=10}
matched_df2_wide %>% filter(Term.Year > 2016)  %>% 
  filter(Academic.Subject == "Biological Sciences") %>%
ggplot(aes(y=urm_gap, x=SI.Attended, color = Class.Subject.Number.Section, 
           group = Random.Course.ID)) + 
  geom_point(aes(size=n)) + geom_line()  + geom_hline(yintercept=0)  +
  ylab("Non-URM - URM")  + xlab("SI Attended") +  guides(size = "none") + labs(color = "Subject and Section") +
  ggtitle("URM gap among matched students who went to SI \n vs those who did not go to SI") + 
  facet_wrap(~Term.Year + Term.Type)
ggsave("poster/pictures/urm_gap_bio.png")
```


```{r}
# Identify courses with 20 or more students (10 matched pairs per class)
matched_courses <- matched_df1 %>% group_by(Random.Course.ID) %>%
  tally() %>% filter(n>=20) %>% droplevels()

matched_df2 <- matched_df1 %>% 
  filter(Random.Course.ID %in% matched_courses$Random.Course.ID) %>%
  group_by(Random.Course.ID, First.Generation.Flag, SI.Attended,
           Term.Year, Academic.Subject) %>%
  summarise(avg.grade = mean(grade), 
            n=n()) 

matched_df2_wide <- matched_df2 %>% 
  pivot_wider(id_cols = c(Random.Course.ID, SI.Attended, Term.Year, Academic.Subject), 
              names_from = First.Generation.Flag, 
              values_from = c(avg.grade, n)) %>%
  mutate(fg_gap = avg.grade_N - avg.grade_Y, 
         n = pmin(n_N, n_Y)) 

ggplot(matched_df2_wide, 
       aes(y=fg_gap, x=SI.Attended, color = Random.Course.ID, 
           group = Random.Course.ID)) + 
  geom_point(aes(size=n)) + geom_line()  + geom_hline(yintercept=0) + 
  geom_text(aes(x=.75, y=.25, label = "Non-FG > FG"), col="red") +
  geom_text(aes(x=.75, y=-.25, label = "Non-FG < FG"), col="green3") +
  ylab("Non-FG - FG") + 
  ggtitle("FG gap among matched students who went to SI vs those who did not go to SI")

matched_df2_wide %>% filter(Term.Year > 2016)  %>%
ggplot(aes(y=fg_gap, x=SI.Attended, color = Academic.Subject, 
           group = Random.Course.ID)) + 
  geom_point(aes(size=n)) + geom_line()  + geom_hline(yintercept=0) + 
  geom_text(aes(x=.75, y=.25, label = "Non-FG > FG"), col="red") +
  geom_text(aes(x=.75, y=-.25, label = "Non-FG < FG"), col="green3") +
  ylab("Non-FG - FG")  +
  ggtitle("FG gap among matched students who went to SI vs those who did not go to SI") + 
  facet_wrap(~Term.Year)
```




* At the course-section level - possibly too small sample sizes here. Sometimes within a class, only 2 students who are a FG who didn't go to SI. We also didn't match on FG. 
* Size of dots is the min size of either FG or NonFG. 

```{r}
matched_df2[33:40,]
```

### Overall (not course level) gap
```{r}
ggplot(matched_df1, aes(y=grade, x = First.Generation.Flag)) + 
  geom_violin(alpha=.5) + geom_jitter(alpha=.5, width=.2, height=.1) +
   stat_summary(fun.data = "mean_cl_boot", colour = "red", size = 1) + 
  facet_wrap(~SI.Attended) 

matched_df1 %>% ggline(x="First.Generation.Flag", y = "grade", 
                       facet.by = "SI.Attended", add=c("mean_se")) + 
  ggtitle("SI helps both grops equally")

```

### Assess treatment affect on matched sample

The statistical quantity of interest is the causal effect of the treatment (`SI.Attended`) on grade in the class: (`grade`). The other variables are pre-treatment covariates. In particular, the analysis is concerned with the marginal, total effect of the treatment for those who actually received the treatment. (ATT) (copy/paste/edit Greifer))_

```{r}
# Following https://cloud.r-project.org/web/packages/MatchIt/vignettes/MatchIt.html
fit1 <- lm(grade ~ SI.Attended + HS.GPA + Student.Orientation.Flag + Academic.Level +
     Major.1.STEM.Flag + First.Generation.Flag + IPEDS.Ethnicity.URM.Non.URM + 
     Gender.Code, 
   data=matched_df1, weights = weights)

coeftest(fit1, vcov. = vcovCL, cluster = ~subclass)
# it is important not to interpret the coefficients and tests of the other covariates in the outcome model. These are not causal effects and their estimates may be severely confounded.
```

The estimated effect was (0.51, SE=.076, p<.0001), indicating that the average effect of attending SI 0.51 grade points higher compared to those that did not go to SI. 

Results still pending further filtering such as on FTF only. And possibly including more years. 

### Strict matching
```{r}
matched_df1.strict <- match.data(matching.strict) %>% 
  rename(grade = Student.Class.Grade.Point.per.Unit)

fit.strict <- lm(grade ~ SI.Attended + HS.GPA + Student.Orientation.Flag + Academic.Level +
     Major.1.STEM.Flag + First.Generation.Flag + IPEDS.Ethnicity.URM.Non.URM + 
     Gender.Code, 
   data=matched_df1.strict, weights = weights)

coeftest(fit.strict, vcov. = vcovCL, cluster = ~subclass)
```

* Hoooooly hell. A near full grade point higher? Too good to be true. What am I missing??

```{r}
ggplot(matched_df1.strict, aes(y=grade, x = First.Generation.Flag)) + 
  geom_violin(alpha=.5) + geom_jitter(alpha=.5, width=.2, height=.1) +
   stat_summary(fun.data = "mean_cl_boot", colour = "red", size = 1) + 
  facet_wrap(~SI.Attended) 

matched_df1.strict %>% ggline(x="First.Generation.Flag", y = "grade", 
                       facet.by = "SI.Attended", add=c("jitter", "mean_se")) 
```

So few FG people?
```{r}
table(matched_df1.strict$SI.Attended, matched_df1.strict$First.Generation.Flag)
```

What about URM?
```{r}
ggplot(matched_df1.strict, aes(y=grade, x = IPEDS.Ethnicity.URM.Non.URM)) + 
  geom_violin(alpha=.5) + geom_jitter(alpha=.5, width=.2, height=.1) +
   stat_summary(fun.data = "mean_cl_boot", colour = "red", size = 1) + 
  facet_wrap(~SI.Attended) 

matched_df1.strict %>% ggline(x="IPEDS.Ethnicity.URM.Non.URM", y = "grade", 
                       facet.by = "SI.Attended", add=c("jitter", "mean_se")) 

table(matched_df1.strict$SI.Attended, matched_df1.strict$IPEDS.Ethnicity.URM.Non.URM)
```

what is the unmodified gap within SI?
```{r}
matched_df1.strict %>% filter(SI.Attended==1) %>% 
  t.test(grade~IPEDS.Ethnicity.URM.Non.URM, data=.)
matched_df1.strict %>% filter(SI.Attended==0) %>% 
  t.test(grade~IPEDS.Ethnicity.URM.Non.URM, data=.)
```


-----

# Does SI reduce the DWF gap
Create dwf variable. This is where we can modify the data used 

```{r}
dod_data <- data %>% 
  mutate(dwf = ifelse(Student.Class.Grade.Point.per.Unit < 1.7, 1, 0))
```

Bivariate exploration. No additional covarites.
```{r}
dod_data %>% dplyr::group_by(IPEDS.Ethnicity.URM.Non.URM, SI.Attended) %>%
  summarise(x = sum(dwf), 
            n=n(), 
            pct = mean(dwf), pct.lab = percent(pct, accuracy = .1))  %>%
  ggplot(aes(x=SI.Attended, 
             y = pct, 
             fill = as.factor(IPEDS.Ethnicity.URM.Non.URM), 
             label = pct.lab)) + 
  geom_col(position=position_dodge()) + 
  geom_text(position=position_dodge(width = 1))
```

* URM gap for Non-SI: 44.4% - 33.6% = 10.8%
* URM gap for SI: 27.1 - 16.8: 10.3%


### Modeling the Difference of Differences
```{r}
dod_mod <- glm(dwf ~ SI.Attended*IPEDS.Ethnicity.URM.Non.URM, 
               data=dod_data)
coef(dod_mod)
```

**Writing contrasts**

* b0 = Intercept
* b1 = SI
* b2 = URM
* b3 = SI * URM

First I define each of the 4 groups, and the calculate the gap within SI and for no SI. 

```{r}
# URM = No & SI = Yes
URM.no_SI.yes <- matrix(c(1, 1, 0, 0), 1)
# URM = Yes & SI = Yes
URM.yes_SI.yes <- matrix(c(1, 1, 1, 1), 1)
(URM.gap_SI.yes <- URM.no_SI.yes - URM.yes_SI.yes)

# URM = No & SI = No
URM.no_SI.no <- matrix(c(1, 0, 0, 0), 1)
# URM = Yes & SI = No
URM.yes_SI.no <- matrix(c(1, 0, 1, 0), 1)
(URM.gap_SI.no <- URM.no_SI.no - URM.yes_SI.no)
```

difference of differences: URM gap for No SI - URM gap for SI
```{r}
(K <- URM.gap_SI.no - URM.gap_SI.yes)

mind.the.gap <- glht(dod_mod, linfct = K)
summary(mind.the.gap)
confint(mind.the.gap)
```

The difference between URM yes and URM does not significantly differ between SI yes and SI no.