---
title: 'Student Level Analysis: SI on One Year Retention'
author: "Faith Fatchen"
date: "4/22/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE )
library(knitr)
library(tidyverse)
library(readxl)
library(magrittr) #Allows %<>% notation to update lhs object with resulting value
library(naniar)
library(Hmisc) # %nin%
library(stargazer)
library(tinytex)
library(ggplot2)
library(dotwhisker)
library(broom)
library(multcomp)

d<-read_csv("data/student_analysis_dataset.csv")
d[sapply(d, is.character)] <- lapply(d[sapply(d, is.character)], as.factor)
d$Term.Year<-factor(d$Term.Year)
d$Cohort.Term.Year<-factor(d$Cohort.Term.Year)
d<-rename(d, URM = IPEDS.Ethnicity.URM.Non.URM)
d$URM.First.Gen<-factor(ifelse(d$URM=="URM" & d$First.Generation.Flag == "Y", 1, 0 )) #new var, student who is urm and first gen
d$One.Year.Retention<-factor(d$One.Year.Retention)

```


```{r eval=FALSE, include=FALSE}
#DATA WORK
# Student Program Dataset
program <- read.csv("data/Student Program.csv")
program.clean <- filter(program, Term.Year >= 2016) %>% # SI classes start in 2016
  mutate(Term.Type = factor(Term.Type), .keep = "unused") %>%
  filter(Term.Type %nin% c("Summer", "Winter")) %>%
  dplyr::select(Term.Year, Random.Student.ID, Gender.Code, IPEDS.Ethnicity,
         IPEDS.Ethnicity.URM.Non.URM, First.Generation.Flag, Academic.Level,
         Academic.Program, Major.1.STEM.Flag, Major.1.College,
         Major.2.STEM.Flag, Major.2.College, Entry.Enrollment.Type,
         Academic.Standing.Status)

program.clean$Random.Student.ID <- factor(program.clean$Random.Student.ID)
# Remove duplicate rows
program.clean <- program.clean[order(program.clean$Random.Student.ID),]
program.clean <- program.clean[!duplicated(program.clean$Random.Student.ID),]

# Student Profile Dataset
profile <- read.csv("data/Student Profile Metric.csv")
profile <- filter(profile, Cohort.Term.Year >= 2016) %>%
  mutate(Random.Student.ID = factor(Random.Student.ID), .keep = "unused") %>%
  filter(Cohort.Student.Enrollment.Type == "First-Time Freshman")%>%
  dplyr::select(Cohort.Term, Cohort.Term.Year, Random.Student.ID, Degree.Term,
         Full.Time.Part.Time.Code, Cohort.Student.Enrollment.Type, HS.GPA.Group,
         HS.GPA, Transfer.GPA.Group, Transfer.GPA, Cohort.Time.to.Degree.Year,One.Year.Retention,
         Student.Orientation.Flag)
program.clean<-filter(program.clean, Random.Student.ID %in% levels(profile$Random.Student.ID))

student_profiles <- left_join(program.clean, profile, by = "Random.Student.ID")
student_profiles <- rename(student_profiles, Enrollment.Term = Cohort.Term)

profile<- student_profiles %>% 
   dplyr::select(c("Random.Student.ID",
           "IPEDS.Ethnicity",
           "IPEDS.Ethnicity.URM.Non.URM",
           "Gender.Code",
           "First.Generation.Flag",
           "Academic.Level",
           "Academic.Program",
           "Major.1.STEM.Flag",
           "Major.1.College",
           "Entry.Enrollment.Type",
           "Academic.Standing.Status",
           "Full.Time.Part.Time.Code",
           "HS.GPA",
           "Enrollment.Term",
           "Cohort.Term.Year",
           "One.Year.Retention",
           "Student.Orientation.Flag"))

grades <- read.csv("data/grades_SI_classes.csv")  %>%
  dplyr::select(c("Random.Course.ID",
           "Term",
           "Term.Year",
           "Term.Type",
           "Student.Class.Unit.Attempted",
           "Student.Class.Unit.Passed",
           "Student.Class.Grade.Point.per.Unit",
           "Student.Class.Official.Grade",
           "Random.Student.ID",
           "Grade.DFW.Count",
           "SI.Visit.Num",
           "SI.Attended"))

courses <- read.csv("data/Course Detail.csv") %>%
  dplyr::select(c("Random.Course.ID",
           "Term",
           "Term.Year",
           "Term.Type",
           "Class.Subject.Number.Section",
           "Course.Subject.and.Number",
           "Academic.Subject",
           "Academic.Subject.Code",
           "Course.Catalog.Number",
           "Class.Learning.Mode",
           "Instruction.Mode",
           "Inst.MD.Persn.Chcoflx.Onl.Othr",
           "Inst.MD.Persn.Onl.Othr",
           "Course.Type",
           "Class.Status",
           "Course.Fee.Exist.Flag",
           "Combined.Section.Group.ID",
           "Writing.Course.Flag",
           "GE.Class.Flag",
           "GE.Foundation.Course",
           "GE.Foundation.Course.Code",
           "GE.Advanced.Course.Substitution.Flag",
           "Honors.Class.Flag"))

grades$Random.Student.ID<-factor(grades$Random.Student.ID)
# Merge the data sets
d <- grades %>% left_join(profile) %>% left_join(courses) %>% drop_na(c(HS.GPA, 
                                                                            Student.Orientation.Flag,
                                                                            Major.1.STEM.Flag,
                                                                            Full.Time.Part.Time.Code,
                                                                            Academic.Program))
                                                                     
write.csv(d, "data/student_analysis_dataset.csv")
```


## Initial Model

```{r}
fit1<-glm(One.Year.Retention ~ Cohort.Term.Year + SI.Attended + Student.Class.Unit.Attempted + Student.Class.Unit.Passed +  URM + Gender.Code + First.Generation.Flag + Academic.Level + Major.1.STEM.Flag + Academic.Standing.Status + HS.GPA + Inst.MD.Persn.Onl.Othr + Course.Fee.Exist.Flag + GE.Class.Flag, family = binomial(link = "logit"), data = d)
summary(fit1)
```

```{r}
m1<-tidy(fit1) %>%
  filter(term == "SI.Attended" | term == "URMURM" | term == "First.Generation.FlagY")
dwplot(m1) +
    geom_vline(xintercept = 0, colour = "grey60", linetype = 2)
```

```{r}
m2<-tidy(fit1) %>%
  filter(term == "Cohort.Term.Year2019" | term == "Cohort.Term.Year2020" | term ==  "Cohort.Term.Year2021")

dwplot(m2) +
    geom_vline(xintercept = 0, colour = "grey60", linetype = 2)

```

```{r}
fit2<-glm(One.Year.Retention ~ Cohort.Term.Year + SI.Visit.Num + Student.Class.Unit.Attempted + Student.Class.Unit.Passed +  URM + Gender.Code + First.Generation.Flag + Academic.Level + Major.1.STEM.Flag + Academic.Standing.Status + HS.GPA + Inst.MD.Persn.Onl.Othr + Course.Fee.Exist.Flag + GE.Class.Flag, family = binomial(link = "logit"), data = d)
summary(fit2)
```

```{r}
m3<-tidy(fit2) %>%
  filter(term == "SI.Visit.Num" | term == "URMURM" | term == "First.Generation.FlagY")
dwplot(m3) +
    geom_vline(xintercept = 0, colour = "grey60", linetype = 2)
```


## With some interactions

```{r}
fit3<-glm(One.Year.Retention~ Cohort.Term.Year + SI.Attended + Student.Class.Unit.Attempted + Student.Class.Unit.Passed +  URM + Gender.Code + First.Generation.Flag + Academic.Level + Major.1.STEM.Flag + Academic.Standing.Status + HS.GPA + Inst.MD.Persn.Onl.Othr + Course.Fee.Exist.Flag + GE.Class.Flag + SI.Attended:URM +SI.Attended:First.Generation.Flag, family = binomial(link = "logit"), data = d)
summary(fit3)
```

```{r}
m4<-tidy(fit3)%>%
  filter(term == "SI.Attended" | term == "SI.Attended:URMURM" | term == "SI.Attended:First.Generation.FlagY")
dwplot(m4) +
    geom_vline(xintercept = 0, colour = "grey60", linetype = 2)
```


```{r}
fit4<-glm(One.Year.Retention~ Cohort.Term.Year + SI.Visit.Num + Student.Class.Unit.Attempted + Student.Class.Unit.Passed +  URM + Gender.Code + First.Generation.Flag + Academic.Level + Major.1.STEM.Flag + Academic.Standing.Status + HS.GPA + Inst.MD.Persn.Onl.Othr + Course.Fee.Exist.Flag + GE.Class.Flag + SI.Visit.Num:URM +SI.Visit.Num:First.Generation.Flag, family = binomial(link = "logit"), data = d)
summary(fit4)
```



```{r}
m5<-tidy(fit4)%>%
  filter(term == "SI.Visit.Num" | term == "SI.Visit.Num:URMURM" | term == "SI.Visit.Num:First.Generation.FlagY")
dwplot(m5) +
    geom_vline(xintercept = 0, colour = "grey60", linetype = 2)
```

## DoD URM

```{r}
dod_data_urm <- d %>% group_by(URM, SI.Attended)

```

### Modeling the Difference of Differences: URM

```{r}
dod_mod_urm <- glm(One.Year.Retention ~ SI.Attended*URM,  family = binomial(link = "logit"), data=dod_data_urm)
coef(dod_mod_urm)

```

**Writing contrasts**

* b0 = Intercept
* b1 = SI
* b2 = URM
* b3 = SI * URM

First I define each of the 4 groups, and the calculate the gap within SI and for no SI. 

```{r}
# URM = No & SI = Yes
URM.no_SI.yes <- matrix(c(1, 1, 0, 0), 1)
# URM = Yes & SI = Yes
URM.yes_SI.yes <- matrix(c(1, 1, 1, 1), 1)
(URM.gap_SI.yes <- URM.no_SI.yes - URM.yes_SI.yes)

# URM = No & SI = No
URM.no_SI.no <- matrix(c(1, 0, 0, 0), 1)
# URM = Yes & SI = No
URM.yes_SI.no <- matrix(c(1, 0, 1, 0), 1)
(URM.gap_SI.no <- URM.no_SI.no - URM.yes_SI.no)
```

difference of differences: URM gap for No SI - URM gap for SI
```{r}
(K <- URM.gap_SI.no - URM.gap_SI.yes)

mind.the.gap.urm <- glht(dod_mod_urm, linfct = K)
summary(mind.the.gap.urm)
confint(mind.the.gap.urm)
```

The difference between URM yes and URM does not significantly differ between SI yes and SI no.

## DoD First gen

```{r}
dod_data_first <- d %>% group_by(First.Generation.Flag, SI.Attended)

```

### Modeling the Difference of Differences: URM

```{r}
dod_mod_first <- glm(One.Year.Retention ~ SI.Attended*First.Generation.Flag,  family = binomial(link = "logit"), data=dod_data_first)
coef(dod_mod_first)

```

**Writing contrasts**
(Same as URM)
* b0 = Intercept
* b1 = SI
* b2 = First
* b3 = SI * First

First I define each of the 4 groups, and the calculate the gap within SI and for no SI. 

```{r}
# First = No & SI = Yes
First.no_SI.yes <- matrix(c(1, 1, 0, 0), 1)
# First = Yes & SI = Yes
First.yes_SI.yes <- matrix(c(1, 1, 1, 1), 1)
(First.gap_SI.yes <- First.no_SI.yes - First.yes_SI.yes)

# First = No & SI = No
First.no_SI.no <- matrix(c(1, 0, 0, 0), 1)
# First = Yes & SI = No
First.yes_SI.no <- matrix(c(1, 0, 1, 0), 1)
(First.gap_SI.no <- First.no_SI.no - First.yes_SI.no)
```

difference of differences: First gap for No SI - First gap for SI
```{r}
(L <- First.gap_SI.no - First.gap_SI.yes)

mind.the.gap.first <- glht(dod_mod_first, linfct = L)
summary(mind.the.gap.first)
confint(mind.the.gap.first)
```

The difference between First yes and First does not significantly differ between SI yes and SI no.

## Plots

```{r}
facet_names<-c(`0` = "Did Not Attend SI", `1` = "Attended SI")
ggplot(dod_data_urm) +
  geom_bar(aes(x=One.Year.Retention, fill=URM),
           position = "dodge") +
  facet_wrap(~SI.Attended, labeller = as_labeller(facet_names)) + xlab("One Year Retention")+ scale_fill_discrete(name="Underrepresented\nMinority(URM)") + ylab("")
```

```{r}
facet_names<-c(`0` = "Did Not Attend SI", `1` = "Attended SI")
ggplot(dod_data_first) +
  geom_bar(aes(x=One.Year.Retention, fill=First.Generation.Flag),
           position = "dodge") +
  facet_wrap(~SI.Attended, labeller = as_labeller(facet_names)) + xlab("One Year Retention")+ scale_fill_discrete(name="First\nGen") + ylab("")
```










